Near Optimal Behavior with Approximate State Abstraction

### 2016-02-03 paper/example_domains.tex
Updated visuals for Minefield and combined figures into single SUPER figure.
Discussion of results now written. Still need to explain plots.

\section{Example Domains}

\enote{Note somewhere why we selected these MDPs -- b/c a wide variety}

To demonstrate the efficacy of these approximate abstraction functions, we test functions in the $\ep{Q^*}$ abstraction family on a variety of domains. We compare the number of states in the abstract MDP to different values of $\epsilon$, as well as the value difference between the optimal policy in the abstract MDP compared to the ground MDP. More specifically, we relate $\epsilon$ to:
\begin{equation}
|V^{\pi_G^*}(s_{init}) - V^{\pi_A^*}(s_{init})|
\end{equation}
For different values of $\epsilon$.

% Description of visuals.
Our code base provides implementations to abstract, visualize, and evaluate an arbitrary MDP to promote further investigation into approximate abstraction. The included visuals of compressed MDPs always converts MDPs into a graph so that an arbitrary MDP may be visualized when abstracted. We use the graph visualization library GraphStream~\cite{graphstream}. In our visuals, a grey circle indicates a state, while the colored arrows indicate transitions. The thickness of the arrow indicates how much reward is associated with that transition. In the ground MDPs, states are labeled with a number. In some abstract MDPs, we indicate which ground states were collapsed to each abstract state (some are visually unwieldy and so were ignored).

\dnote{One thing we want to add is a description of how diverse these domains are!}
\dnote{Also put in a note that we're using a uniform weighting scheme}

@@

% Minefield Compression Visuals
\begin{figure}
\subfigure[Ground Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/ground_minefield.png}}
\subfigure[Abstract Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/abs_minefield.png}}
\label{fig:minefield-visual}
\caption{Comparison of Original Minefield MDP and Abstract MDP, under an $\epQ$ abstraction, with $\epsilon=1.1$}
\end{figure}

>>>>

\section{Example Domains}
To demonstrate the efficacy of these approximate abstraction functions, we test functions in the $\ep{Q^*}$ abstraction family on four domains -- NChain, Taxi, Minefield and Random MDP, which we dub ``Random". These domains were selected for the variety they offer -- Nchain is relatively simple domain, Taxi is goal-based and hierarchical in nature, Minefield is stochastic and Random MDP has a high number of near-optimal behavior.

% Description of visuals.
Our code base provides implementations to abstract, visualize, and evaluate an arbitrary MDP to promote further investigation into approximate abstraction. The included visuals of compressed MDPs always converts MDPs into a graph so that an arbitrary MDP may be visualized when abstracted. We use the graph visualization library GraphStream~\cite{graphstream}. In our visuals, a grey circle indicates a state, while the colored arrows indicate transitions. The thickness of the arrow indicates how much reward is associated with that transition. In the ground MDPs, states are labeled with a number. In some abstract MDPs, we indicate which ground states were collapsed to each abstract state by labelling the abstract states with their constituent ground states (some are visually unwieldy and so were ignored).

@@

% Minefield Compression Visuals
\begin{figure}
\subfigure[Ground Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/ground_minefield.png}}
\subfigure[Abstract Minefield]{
\includegraphics[width=0.48\columnwidth]{figures/abs_minefield.png}}
\label{fig:minefield-visual}
\caption{Comparison of Original Minefield MDP and Abstract MDP, under an $\epQ$ abstraction, with $\epsilon=.8$}
\end{figure}

### 2016-02-03 paper/introduction.tex
Discussion of results now written. Still need to explain plots.

% Approximate abstraction
In this work, we demonstrate that aggregation of states on the basis of various approximate criteria incur only bounded error in the resulting policy. Relaxing the aggregation criteria from equality of quantities to $\varepsilon$-closeness offers two benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge that we expect a planning or learning algorithm to approximate without fully solving the \ac{MDP}. Second, because the state-aggregation criteria are relaxed to approximate equality, methods that employ approximate equality are able to tune the aggressiveness of state aggregation all the while incurring bounded error.

>>>>

% Approximate abstraction
The thesis of this work is that that aggregation of states on the basis of various approximate criteria incur only bounded error in the resulting policy. Relaxing the aggregation criteria from equality of quantities to $\varepsilon$-closeness offers two benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge that we expect a planning or learning algorithm to approximate without fully solving the \ac{MDP}. Second, because the state-aggregation criteria are relaxed to approximate equality, methods that employ approximate equality are able to tune the aggressiveness of state aggregation all the while incurring bounded error.

### 2016-02-03 paper/results_discussion.tex
Discussion of results now written. Still need to explain plots.

Here we discuss the results shown in states blah blah \dnote{Description of this section}.

\begin{itemize}
\item For $\epsilon > 0$, there are extremely useful abstractions to be made. Want to emphasize that with approximate knowledge of X, can still abstract effectively (point out Minefield + NChain both retain a very good policy from abstract MDP). This hits the learnable point.
\item Explain goal based/Taxi (regenerate plot for different scale of epsilon)
\item Possibly talk about stochasticity (low priority)
\item Explain Random behavior, smooth interpolation, mention randomness, etc.
\item Describe plots a bit (what is the y-axis?)
\end{itemize}

% Figure: Epsilon vs. #States for all three sample domains.
\begin{figure}
\centering

@@

\subfigure[RandomStates]{
\includegraphics[width=0.46\columnwidth]{figures/random_epsilon_vs_num_abstract_states.pdf}
}
\subfigure[RandomVal]{
\includegraphics[width=0.46\columnwidth]{figures/random_epsilon_vs_value_of_abstract_policy_init_ground_state.pdf}
}

\label{fig:main results}
\caption{$\epsilon$ vs. Num States and $\epsilon$ vs. Abstract Policy Value}
\end{figure}

>>>>

%Explanation of what abstraction is used
We compare the number of states in the abstract MDP to different values of $\epsilon$, as well as the value difference between the optimal policy in the abstract MDP compared to the ground MDP. More specifically, we relate $\epsilon$ to:
\begin{equation}
|V^{\pi_G^*}(s_{init}) - V^{\pi_A^*}(s_{init})|
\end{equation}
For different values of $\epsilon$.

\dnote{Also put in a note that we're using a uniform weighting scheme}
\enote{Also put in a note how we're randomly choosing a Q*}
%Explain plots
Here we discuss the results shown in states blah blah \dnote{Description of this section}.

%Discuss results
Our empirical results corroborate our thesis -- that approximate state abstractions can form the basis of learnable and useful state aggregation criteria. In both NChain and Minefield, we observe that despite increasing $\epsilon$ to reduce the number of states which must be planned over, optimal behavior is either fully maintained (NChain) or optimal behavior is very nearly maintained (Minefield). Similarly for Taxi and $\epsilon$ between $.02$ and $.025$ we observe a reduction in the number of states that must be planned over while value is fully maintained. After $.025$, increased reduction in state space size comes at a cost of value. Lastly, as $\epsilon$ is increased Random smoothly reduces the number of abstract states which must be planned over with a corresponding cost in the value of the derived policy.

Our experimental results also demonstrate a noteworthy contrast between approximate state abstraction on goal-based \acp{MDP} and random \acp{MDP}. Taxi exhibits relative stability in state space size and behavior for epsilon up to $.02$ until both precipitously fall off. We attribute the sudden fall off in number of abstract states and value of the derived policy in Taxi to its goal-based nature; once certain information critical for optimal behavior is lost in the state aggregation, solving the goal -- and so acquiring any reward -- is rendered entirely unattainable. Conversely, in Random a great deal of near-optimal policies are available to the agent. Thus, even the information for optimal behavior is lost with an increase in $\epsilon$ there are a great deal of near-optimal policies available to the agent which are not lost in the state aggregation.

% Figure: Epsilon vs. #States and Epsilon vs Abstract Pol. Value for all four sample domains.
\begin{figure}
\centering

@@

\subfigure[Random]{
\includegraphics[width=0.46\columnwidth]{figures/random_epsilon_vs_num_abstract_states.pdf}
}
\subfigure[Random]{
\includegraphics[width=0.46\columnwidth]{figures/random_epsilon_vs_value_of_abstract_policy_init_ground_state.pdf}
}

\label{fig:main_empirical_results}
\caption{$\epsilon$ vs. Num States and $\epsilon$ vs. Abstract Policy Value}
\end{figure}
